Wetting of the tarsal adhesive fluid controls underwater adhesion in ladybug beetles

### 2021-07-14 source/jeb_tex/manuscript.tex
changed bibiliography style file to match reference formatting
fixed citations

While most of the studies on insect adhesion focused on terrestrial species, underwater insect attachment is much rarer and has been relatively unexplored. Some aquatic insects like diving beetles \citep{Chen:2014} or midge larva \citep{Kang:2020} use suction cups to adhere to surfaces. However, underwater adhesion mediated by secreted liquids require the displacement of the water at the interface first and a spreading of the fluid on the substrate. One relatively simple approach is to use an air bubble around the adhesive organs similar to the air bubbles many secondary aquatic insects and spiders carry on their body for breathing underwater \citep{Seymour:2013}. This has been shown in a recent study by \citep{RN87} that female terrestrial green dock beetles \emph{Gastrophysa viridula} can attach quite well to surfaces underwater by using such an air bubble. Their naturally hydrophobic tarsal hairs trap the bubble around the pads when being submerged underwater, which de-wets the surface on contact. It has been hypothesised that a combination of capillary forces due the air bubble and hair secretions within the de-wetted area results in its adhesion underwater. However, it remained unclear if an air bubble is necessary for adhesion and what, if any, contribution it has to the adhesive force. The oily tarsal adhesive fluid found in insects alone might be sufficient in creating the necessary capillary adhesion even without a bubble, given that the fluid remains on the hair tips when submerged. In ladybug beetles, the tip of each seta secretes approximately one femtoliter of tarsal adhesive fluid by each step \citep{RN108}. The fluid's chemical composition in green dock beetles was identified to be an oil-containing mixture of mostly long chain hydrocarbons\citep{RN96} with traces
of triglycerides, fatty acids and cholesterol in ladybirds \emph{Hemisphaerota cyanea} \citep{RN221} and \emph{Epilachna vingtipunctuata} \citep{RN222}, rendering it immiscible with water.

@@

To some extent, the findings could be extended to other animals relying on oily
secretions for adhesion. For example, ants are known to possess smooth adhesive pads which secrete a fluid containing oily substances \citep{RN201}.
It has been reported that some ants show similar adhesion on hydrophobic substrates under wet and dry conditions \citep{RN213},
similar to what we see in a ladybug. This observation can again be explained by a capillary model as before. Recent experiments on
geckos revealed that they can attach well to fluoropolymer substrates (such as PTFE)
underwater while they show little adhesion to the same substrate in
air\citep{RN199,RN15}. Geckos are thought to rely on van der Waals
forces via dry contact with the substrate \citep{RN202}, although
recent observations of phospholipid footprints left behind walking
geckos \citep{RN205} could change that picture. Since geckos adhere
poorly to PTFE (surface energy $\sim$ 20 mN m\protect\textsuperscript{-1}) one can
speculate that the phospholipid material has a higher surface energy,
and consequently makes a higher contact angle with PTFE in air. Let
us assume the phosopholipid substance to be a fluid similar to oil 
with $\gamma_{fa}$= 30 mN m\protect\textsuperscript{-1} and $\gamma_{fw}$= 42 mN m\protect\textsuperscript{-1} such that
its contact angle with PTFE is 80\textdegree . Equation \ref{eq:theta_fw}
then gives us an underwater contact angle of 70\textdegree{} for the
phospholipid fluid. Thus, on a PTFE surface, the capillary bridge
model can predict a higher adhesion underwater than in air due to
its lower contact angle and higher interfacial energy underwater.
Based on similar assumptions, we predict the net adhesion force for
the gecko on different substrates (Figure \ref{fig:Comparison-with-model}).
The adhesion force predictions are in good qualitative agreement with
the whole animal experimental shear force values reported for the
gecko, with the trend of higher adhesion in air than underwater for
glass, similar adhesion in air and underwater for PMMA/OTS-SAM and
lower adhesion in air than underwater for PTFE. We, thus, propose
that the underwater experiments performed on geckos \citep{RN199,RN15}
indicate a capillary contribution to gecko adhesion. Previous studies on gecko adhesion have attributed capillary effects to be a result of water monolayers adsorbed from ambient humid air onto the spatuale hair tips \citep{RN197, RN203, RN218}.
We however emphasize that capillary contribution in gecko adhesion could rather be a result of its setal phospholipid layer rather than water.
The previously reported influence of humidity on gecko adhesion \citep{RN197} could possibly be an effect of change in surface tension of the oily phospholipid layer at different humidity, which will in-turn influence the capillary adhesion force. We suggest performing single seta adhesion force tests similar to \citep{RN202}
using a hydrophilic and fluorinated probe in air and underwater conditions. 
If the fluorinated probe shows higher adhesion underwater than in air on a single seta, this
would confirm the role of capillary contributions due to an oil-like phospholipid layer to gecko adhesion. Further work is also necessary to validate if the phospholipid layer in geckos' toes indeed has fluid-like properties.

\begin{figure}
\centering
\includegraphics[width=3.5in]{Figure-9-Gecko_comparison}\caption{\label{fig:Comparison-with-model}Whole animal adhesion force of geckos
on various substrates. Experimental shear adhesion values are reproduced
from \citep{RN15}. Normal adhesion forces for each gecko
toe are theoretically estimated from the capillary bridge model, with
hair diameter = 400 nm, toe diameter = 4 mm, phospholipid fluid volume
= 4.19x10\protect\textsuperscript{-3} fL and 10\% hair coverage.
\textquotedblleft\emph{Underwater: no bubble}\textquotedblright{}
contact mode is assumed for the \textquotedblleft Wet\textquotedblright{}
case. Net adhesion force is calculated by assuming 5 toes on each
leg and 4 legs in total on a gecko. Interfacial tension of the phospholipid
layer (PL) in air and water are assumed to be 30 mN m\protect\textsuperscript{-1} and 42 mN m\protect\textsuperscript{-1}
respectively. PL contact angles with glass, PMMA, OTS-SAM and PTFE
are assumed to be 6\textdegree , 10\textdegree , 20\textdegree{} and
80\textdegree{} respectively. The corresponding water contact angles
are 50\textdegree , 85\textdegree , 94\textdegree{} and 97\textdegree{}
respectively, as reported in \citep{RN15}.}
\end{figure}

We have so far limited our analysis to only smooth substrates. Of course insects
have to cope with all kinds of surfaces including rough ones.
Previous studies \citep{RN136} have shown that substrate roughness
is a more dominant parameter than substrate chemistry in controlling
ladybug beetle traction force. Future work will explore how roughness can impact
the net capillary force also in wet and submerged conditions. In our study, we have only considered normal adhesive forces, but insects like beetles in general rely on friction or shear forces during locomotion. Friction force usually correlates directly with the normal force, which is probably why previously reported shear adhesion forces of the dock beetle \citep{RN87} follow a similar qualitative trend as our normal adhesion force measurements on the ladybug beetle in both air and underwater conditions. However, the details of the interplay between friction and normal adhesion forces in animals is an open question and is beyond the scope of this paper.

@@

%\begin{thebibliography}{}
%
%\bibitem{RN24}
%{\bf Arzt, E., S.~Gorb, and R.~Spolenak.}  2003.
%\textit{ Proc Natl Acad Sci U S A}  {\bf 100} (19):10603--6.
%
%\bibitem{RN221}
%{\bf Attygalle, A.~B., D.~J. Aneshansley, J.~Meinwald, and T.~Eisner.}  2000.
%\textit{ Zoology}  {\bf 103} (1/2):1--6.
%
%\bibitem{RN202}
%{\bf Autumn, K., Metin Sitti, Yiching~A. Liang, Anne~M. Peattie, Wendy~R.
%  Hansen, Simon Sponberg, Thomas~W. Kenny, Ronald Fearing, Jacob~N.
%  Israelachvili, and Robert~J. Full.}  2002.
%\textit{ Proceedings of the National Academy of Sciences}  {\bf 99} (19):12252.
%
%\bibitem{RN206}
%{\bf Brakke, K.~A.}  1992.
%\textit{ Experiment. Math.} {\bf 1} (2):141--165.
%
%\bibitem{RN19}
%{\bf Bullock, J.~M., and W.~Federle.}  2009.
%\textit{ J Exp Bio}  {\bf 212} (Pt 12):1876--88.
%
%\bibitem{RN79}
%{\bf Bullock, J.~M., and W.~Federle.}  2011.
%\textit{ Naturwissenschaften}  {\bf 98} (5):381--7.
%
%\bibitem{Chen:2014}
%{\bf Chen, Y., Ming-Chih Shih, Ming-Huang Wu, En-Cheng Yang, and Kai-Jung Chi.}
%   2014.
%\textit{ Journal of The Royal Society Interface}  {\bf 11} (97).
%
%\bibitem{Coddington:1991}
%{\bf Coddington, J.~A., and H~W Levi.}  1991.
%\textit{ Annual Review of Ecology and Systematics}  {\bf 22}:565--592.
%
%\bibitem{RN93}
%{\bf De~Souza, E.~J., M.~Brinkmann, C.~Mohrdieck, and E.~Arzt.}  2008.
%\textit{ Langmuir}  {\bf 24} (16):8813--8820.
%
%\bibitem{RN71}
%{\bf Dirks, J.~H.}  2014.
%\textit{ Beilstein J Nanotechnol}  {\bf 5}:1160--6.
%
%\bibitem{RN136}
%{\bf England, M.~W., Tomoya Sato, Makoto Yagihashi, Atsushi Hozumi,
%  Stanislav~N. Gorb, and Elena~V. Gorb.}  2016.
%\textit{ Beilstein journal of nanotechnology}  {\bf 7}:1471--1479.
%
%\bibitem{RN20}
%{\bf Federle, W.}  2006.
%\textit{ J Exp Biol}  {\bf 209} (Pt 14):2611--21.
%
%\bibitem{RN201}
%{\bf Federle, W., Mathis Riehle, Adam~S.G. Curtis, and Robert~J. Full.}  2002.
%\textit{ Integrative and Comparative Biology}  {\bf 42} (6):1100--1106.
%
%\bibitem{RN96}
%{\bf Geiselhardt, S.~F., Sven Geiselhardt, and Klaus Peschke.}  2009.
%\textit{ Chemoecology}  {\bf 19} (4):185.
%
%\bibitem{RN145}
%{\bf Heepe, L., Dennis~S. Petersen, Lisa Tölle, Jonas~O. Wolff, and
%  Stanislav~N. Gorb.}  2016.
%\textit{ Applied Physics A}  {\bf 123} (1):34.
%
%\bibitem{RN87}
%{\bf Hosoda, N., and S.~N. Gorb.}  2012.
%\textit{ Proc Biol Sci}  {\bf 279} (1745):4236--42.
%
%\bibitem{RN205}
%{\bf Hsu, P.~Y., Liehui Ge, Xiaopeng Li, Alyssa~Y. Stark, Chrys Wesdemiotis,
%  Peter~H. Niewiarowski, and Ali Dhinojwala.}  2012.
%\textit{ Journal of The Royal Society Interface}  {\bf 9} (69):657--664.
%
%\bibitem{RN197}
%{\bf Huber, G., Hubert Mantz, Ralph Spolenak, Klaus Mecke, Karin Jacobs,
%  Stanislav~N. Gorb, and Eduard Arzt.}  2005.
%\textit{ Proceedings of the National Academy of Sciences of the United States
%  of America}  {\bf 102} (45):16293--16296.
%
%\bibitem{RN222}
%{\bf Ishii, S.}  1987.
%\textit{ Applied Entomology and Zoology}  {\bf 22} (2):222--228.
%
%\bibitem{Kang:2020}
%{\bf Kang, V., Robin~T. White, Simon Chen, and Walter Federle.}  2020.
%\textit{ bioRxiv}  {\bf }.
%
%\bibitem{RN203}
%{\bf Kim, T.~W., and Bharat Bhushan.}  2008.
%\textit{ Journal of The Royal Society Interface}  {\bf 5} (20):319--327.
%
%\bibitem{RN155}
%{\bf Langer, M.~G., J.~P. Ruppersberg, and S.~Gorb.}  2004.
%\textit{ Proceedings of the Royal Society of London. Series B: Biological
%  Sciences}  {\bf 271} (1554):2209--2215.
%
%\bibitem{RN218}
%{\bf Mitchell, C.~T., Cem~Balda Dayan, Dirk-M. Drotlef, Metin Sitti, and
%  Alyssa~Y. Stark.}  2020.
%\textit{ Scientific Reports}  {\bf 10} (1):19748.
%
%\bibitem{RN108}
%{\bf Peisker, H., and Stanislav~N. Gorb.}  2012.
%\textit{ The Journal of Experimental Biology}  {\bf 215} (8):1266--1271.
%
%\bibitem{Seymour:2013}
%{\bf Seymour, R.~S., and Philip G.~D. Matthews.}  2013.
%\textit{ Journal of Experimental Biology}  {\bf 216} (2):164--170.
%
%\bibitem{RN15}
%{\bf Stark, A.~Y., I.~Badge, N.~A. Wucinich, T.~W. Sullivan, P.~H.
%  Niewiarowski, and A.~Dhinojwala.}  2013.
%\textit{ Proc Natl Acad Sci U S A}  {\bf 110} (16):6340--5.
%
%\bibitem{RN199}
%{\bf Stark, A.~Y., Daniel~M. Dryden, Jeffrey Olderman, Kelly~A. Peterson,
%  Peter~H. Niewiarowski, Roger~H. French, and Ali Dhinojwala.}  2015.
%\textit{ Journal of The Royal Society Interface}  {\bf 12} (108):20150464.
%
%\bibitem{RN213}
%{\bf Stark, A.~Y., and Stephen~P. Yanoviak.}  2018.
%\textit{ Royal Society open science}  {\bf 5} (11):181540--181540.
%
%\bibitem{RN59}
%{\bf Stork, N.~E.}  1980.
%\textit{ The Journal of Experimental Biology}  {\bf 88} (1):91.
%
%\bibitem{Williams:1982}
%{\bf Williams, E., and J.A. Peterson.}  1982.
%\textit{ Science}  {\bf 215}:1509--1511.
%
%\end{thebibliography}

%\bibitem{RN198}
%{\bf Hooke, R.}  1665.
%\newblock {\em Micrographia, or, Some physiological descriptions of minute
%  bodies made by magnifying glasses : with observations and inquiries
%  thereupon}.
%\newblock The Royal Society.

>>>>

While most of the studies on insect adhesion focused on terrestrial species, underwater insect attachment is much rarer and has been relatively unexplored. Some aquatic insects like diving beetles \citep{Chen:2014} or midge larva \citep{Kang:2020} use suction cups to adhere to surfaces. However, underwater adhesion mediated by secreted liquids require the displacement of the water at the interface first and a spreading of the fluid on the substrate. One relatively simple approach is to use an air bubble around the adhesive organs similar to the air bubbles many secondary aquatic insects and spiders carry on their body for breathing underwater \citep{Seymour:2013}. This has been shown in a recent study by \citet{RN87} that female terrestrial green dock beetles \emph{Gastrophysa viridula} can attach quite well to surfaces underwater by using such an air bubble. Their naturally hydrophobic tarsal hairs trap the bubble around the pads when being submerged underwater, which de-wets the surface on contact. It has been hypothesised that a combination of capillary forces due the air bubble and hair secretions within the de-wetted area results in its adhesion underwater. However, it remained unclear if an air bubble is necessary for adhesion and what, if any, contribution it has to the adhesive force. The oily tarsal adhesive fluid found in insects alone might be sufficient in creating the necessary capillary adhesion even without a bubble, given that the fluid remains on the hair tips when submerged. In ladybug beetles, the tip of each seta secretes approximately one femtoliter of tarsal adhesive fluid by each step \citep{RN108}. The fluid's chemical composition in green dock beetles was identified to be an oil-containing mixture of mostly long chain hydrocarbons\citep{RN96} with traces
of triglycerides, fatty acids and cholesterol in ladybirds \emph{Hemisphaerota cyanea} \citep{RN221} and \emph{Epilachna vingtipunctuata} \citep{RN222}, rendering it immiscible with water.

@@

To some extent, the findings could be extended to other animals relying on oily
secretions for adhesion. For example, ants are known to possess smooth adhesive pads which secrete a fluid containing oily substances \citep{RN201}.
It has been reported that some ants show similar adhesion on hydrophobic substrates under wet and dry conditions \citep{RN213},
similar to what we see in a ladybug. This observation can again be explained by a capillary model as before. Recent experiments on
geckos revealed that they can attach well to fluoropolymer substrates (such as PTFE)
underwater while they show little adhesion to the same substrate in
air\citep{RN199,RN15}. Geckos are thought to rely on van der Waals
forces via dry contact with the substrate \citep{RN202}, although
recent observations of phospholipid footprints left behind walking
geckos \citep{RN205} could change that picture. Since geckos adhere
poorly to PTFE (surface energy $\sim$ 20 mN m\protect\textsuperscript{-1}) one can
speculate that the phospholipid material has a higher surface energy,
and consequently makes a higher contact angle with PTFE in air. Let
us assume the phosopholipid substance to be a fluid similar to oil 
with $\gamma_{fa}$= 30 mN m\protect\textsuperscript{-1} and $\gamma_{fw}$= 42 mN m\protect\textsuperscript{-1} such that
its contact angle with PTFE is 80\textdegree . Equation \ref{eq:theta_fw}
then gives us an underwater contact angle of 70\textdegree{} for the
phospholipid fluid. Thus, on a PTFE surface, the capillary bridge
model can predict a higher adhesion underwater than in air due to
its lower contact angle and higher interfacial energy underwater.
Based on similar assumptions, we predict the net adhesion force for
the gecko on different substrates (Figure \ref{fig:Comparison-with-model}).
The adhesion force predictions are in good qualitative agreement with
the whole animal experimental shear force values reported for the
gecko, with the trend of higher adhesion in air than underwater for
glass, similar adhesion in air and underwater for PMMA/OTS-SAM and
lower adhesion in air than underwater for PTFE. We, thus, propose
that the underwater experiments performed on geckos \citep{RN199,RN15}
indicate a capillary contribution to gecko adhesion. Previous studies on gecko adhesion have attributed capillary effects to be a result of water monolayers adsorbed from ambient humid air onto the spatuale hair tips \citep{RN197, RN203, RN218}.
We however emphasize that capillary contribution in gecko adhesion could rather be a result of its setal phospholipid layer rather than water.
The previously reported influence of humidity on gecko adhesion \citep{RN197} could possibly be an effect of change in surface tension of the oily phospholipid layer at different humidity, which will in-turn influence the capillary adhesion force. We suggest performing single seta adhesion force tests similar to \citet{RN202}
using a hydrophilic and fluorinated probe in air and underwater conditions. 
If the fluorinated probe shows higher adhesion underwater than in air on a single seta, this
would confirm the role of capillary contributions due to an oil-like phospholipid layer to gecko adhesion. Further work is also necessary to validate if the phospholipid layer in geckos' toes indeed has fluid-like properties.

\begin{figure}
\centering
\includegraphics[width=3.5in]{Figure-9-Gecko_comparison}\caption{\label{fig:Comparison-with-model}Whole animal adhesion force of geckos
on various substrates. Experimental shear adhesion values are reproduced
from \citet{RN15}. Normal adhesion forces for each gecko
toe are theoretically estimated from the capillary bridge model, with
hair diameter = 400 nm, toe diameter = 4 mm, phospholipid fluid volume
= 4.19x10\protect\textsuperscript{-3} fL and 10\% hair coverage.
\textquotedblleft\emph{Underwater: no bubble}\textquotedblright{}
contact mode is assumed for the \textquotedblleft Wet\textquotedblright{}
case. Net adhesion force is calculated by assuming 5 toes on each
leg and 4 legs in total on a gecko. Interfacial tension of the phospholipid
layer (PL) in air and water are assumed to be 30 mN m\protect\textsuperscript{-1} and 42 mN m\protect\textsuperscript{-1}
respectively. PL contact angles with glass, PMMA, OTS-SAM and PTFE
are assumed to be 6\textdegree , 10\textdegree , 20\textdegree{} and
80\textdegree{} respectively. The corresponding water contact angles
are 50\textdegree , 85\textdegree , 94\textdegree{} and 97\textdegree{}
respectively, as reported in \citet{RN15}.}
\end{figure}
%CHANGE REFERENCE FORMAT IN PLOT
We have so far limited our analysis to only smooth substrates. Of course insects
have to cope with all kinds of surfaces including rough ones.
Previous studies \citep{RN136} have shown that substrate roughness
is a more dominant parameter than substrate chemistry in controlling
ladybug beetle traction force. Future work will explore how roughness can impact
the net capillary force also in wet and submerged conditions. In our study, we have only considered normal adhesive forces, but insects like beetles in general rely on friction or shear forces during locomotion. Friction force usually correlates directly with the normal force, which is probably why previously reported shear adhesion forces of the dock beetle \citep{RN87} follow a similar qualitative trend as our normal adhesion force measurements on the ladybug beetle in both air and underwater conditions. However, the details of the interplay between friction and normal adhesion forces in animals is an open question and is beyond the scope of this paper.

@@
